---
title: "06_stats"
format: html
editor: visual
---

```{r}
pacman::p_load(tidyverse, here, janitor)
```

# Stats

**Descriptive**: describe and summarize data

**Inferential**: use a sample data to make inferences about a larger population

## Types of data

**Numeric** (quantative): continuos or discrete

**Categorical** (qualitative): nominal or ordinal

## Measures of center

-   mean,

-   median,

-   mode

```{r}
mean(sleep_data$sleep_hours)
median(sleep_data$sleep_hours)
```

## Measures of spread

-   variance, sd (penalize longer distances), mean absolute deviation (MAD) (dont penalize longer distance)

-   quartiles, quantile, IQR

#### **EXCERCISES**

1.  Load the dataset and explore its structure.

    ```{r}
    data("mtcars")
    ```

    ```{r}
    glimpse(mtcars)
    ```

2.  Compute summary statistics for the `mpg` (miles per gallon) variable.

    ```{r}
    summary(mtcars$mpg)
    ```

3.  Compute the mean, median, and mode of `mpg`.

    ```{r}
    mean(mtcars$mpg)

    ```

    ```{r}
    median(mtcars$mpg)
    ```

4.  Compute the variance, standard deviation, and interquartile range (IQR) for `mpg`.

```{r}
var(mtcars$mpg)
```

```{r}
sd(mtcars$mpg)
```

```{r}
IQR(mtcars$mpg)
```

```{r}
mtcars |> 
  ggplot(aes(x = mpg)) + 
  geom_histogram(bins = 10)
```
