---
title: "VIM_Missing_values"
format: html
editor: visual
---

## 

```{r}
pacman::p_load(tidyverse, 
               visdat, 
               VIM )
```

```{r}
data("sleep")
```

```{r}
visdat::vis_dat(sleep)
```

```{r}
aggr(sleep, plot = FALSE) |> 
  plot(numbers = TRUE, prop = FALSE)
```

The left plot shows the amount of missings for each column in the dataset `sleep` and the right plot shows how often each combination of missings occur. For example, there are 9 rows wich contain a missing in both `NonD` and `Dream`.

```{r}
 sleep[, c("Dream", "Sleep")] |> 
   marginplot()
```

The **red** boxplot on the left shows the distrubution of all values of `Sleep` where `Dream` contains a missing value. The **blue** boxplot on the left shows the distribution of the values of `Sleep` where `Dream` is observed.

## Impute missing values

In order to impute missing values, `VIM` offers a spectrum of imputation methods like `kNN()` (k nearest neighbour), `hotdeck()` and so forth. Those functions can be applied to a `data.frame` and return another `data.frame` where missings are replaced by imputed values.

```{r}
x <- sleep[, c("Dream", "Sleep")]
x_imputed <- kNN(x)
```

```{r}
marginplot(x_imputed, delimiter = "_imp")
```

In this plot three differnt colors are used in the top-right. These colors represent the structure of missings.

-   **brown** points represent values where `Dream` was missing initially

-   **beige** points represent values where `Sleep` was missing initially

-   **black** points represent values where both `Dream` and `Sleep` were missing initially

The `kNN()` method seemingly preserves the correlation between `Dream` and `Sleep`.
